Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2017-09-06 redaction/Presentations/DEGIT_Conference/slides_DEGIT_v1.tex
Nouvelle version des slides

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] Part of the ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\item[$\Rightarrow$] Multiplicative costs:     \footnotesize $pq = \tau \widetilde{p}q$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2015) \vspace{0.1cm}
\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
\item[$\Rightarrow$] Additive costs: \footnotesize $pq = (t+\widetilde{p})q  $
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item The structure (additive vs iceberg) of transport costs, not anecdotal \vspace{0.2cm}
\begin{itemize}
\item[-] Additive costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}
\item[-] The additive structure of trade costs is supported by recent empirical evidence (Irrarazabal et al. 2014, Hummels \& Skiba, 2004)  \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Transport costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}

@@

%}
\end{itemize}
\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' %}
%}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item[(3)] \textbf{In accounting for the time trend of transport costs} \vspace{0.1cm}
\begin{itemize}
\item[-] The importance of excluding composition effects when characterizing the time trend (Hummels, 2007) \vspace{0.1cm}
\item[-] Our contribution: The importance of modelling the additive component of transport costs \vspace{0.1cm}
\item[$\Rightarrow$] The decrease in TC over time is mostly attributable to the decrease in the ``pure'' transport costs,  \vspace{0.1cm}
\item[-] Rather than to trade composition effects  \vspace{0.1cm}
\item[-] A result in sharp contrast with Hummels (2007) \vspace{0.2cm}
\end{itemize}
\end{itemize}
\textbf{All our results: Provide new quantitative evidence about the importance of the additive component in international transport costs}

@@

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, SITC 5-digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
%(Could go up to HS-10 with recent data)\vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] The RHS are at the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 sectors (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 sectors at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, insurance + handling + quantitative freight costs (e.g. not those related to the time value of goods) \vspace{0.1cm}
%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, long time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
\item[-] Have both the import and the export prices  \vspace{0.1cm}
\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive transport costs  \vspace{0.1cm}
\begin{itemize}
\item[$\neq$] Irrarazabal et al. (2015)
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:basis_equation}
\end{equation}
\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also mode (air or vessel)- specific
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\end{itemize}
\footnotesize
\begin{eqnarray*}
\tau_{ik} = \tau_i \times \tau_k &\text{and}& t_{ik} = t_i + t_k
\end{eqnarray*}
\normalsize
\item Additional assumption: All products in a 3-digit sector ($s$) share the same structure of costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{is(k)}$ and $\tau_{is(k)}$ as:
\footnotesize
\begin{equation}
 \tau_{is(k)} = \tau_{i} \times \tau_{s(k)}, \qquad t_{is(k)} = t_{i} + t_{s(k)} \label{eq:specifTC}
 \end{equation}
 \normalsize
\end{itemize}
\end{frame}

\begin{frame}[label=slide_estimation_method]
\begin{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} >1$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The equation of interest (\ref{eq:basis_equation}) becomes:
\footnotesize
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{s(k)} -1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\normalsize
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation*}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
\end{equation*}
\normalsize
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
\end{itemize}
\item A non-linear equation (due to the additive costs)  \vspace{0.1cm}
\item[$\Rightarrow$] Estimation using non-linear squares \hyperlink{app_estimation_method}{\beamergotobutton{More}}\vspace{0.1cm}

\begin{frame}
\frametitle{Empirical specification (3)}
\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
\begin{itemize}
\item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\item[$\Rightarrow$] Estimate \textit{two} equations, for each year over 1974-2013, by transport mode (air/ocean)\vspace{0.1cm}
%\begin{itemize}
\item[(A)] With only ad-valorem costs:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\normalsize
\item[(B)] With additive costs included:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\begin{itemize}
\item[$\ast$] By year and transport mode, $\simeq$ 800 fixed effects to estimate   \vspace{0.1cm}
%\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_result1]
\frametitle{Result 1: Size of additive transport costs \hyperlink{app_result1}{\beamergotobutton{More on this}} }
\begin{itemize}
\item \textbf{International transport costs: A sizeable additive component}  \vspace{0.2cm}
\begin{itemize}
\item[-] 48.2\% of total costs in average for ocean, 42.3\% for air  \vspace{0.1cm}
\item[-] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
\begin{itemize}
\footnotesize{
\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\ast$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.1cm}
\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
\end{itemize}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.  \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Time trends in transport costs \& composition effects }
\begin{itemize}
\item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.2cm}
\item The change in overall transport costs over time: \vspace{0.1cm}
\begin{itemize}
\item[-] Depend on the evolution of per product- per partner costs, \vspace{0.1cm}
\item[-] But also on the composition of trade flows \vspace{0.2cm}
\begin{itemize}
\item[$\star$] Over time, import more goods that are cheaper to transport, and/or from countries with which it is cheaper to trade \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
\item[$\Rightarrow$] What we do, in accordance with Hummels (2007)
\end{itemize}
\end{frame}

\begin{frame} [label = slide_compeffects_strategy]
\frametitle{Eliminating composition effects: Our strategy}
\begin{itemize}
\item Based on our estimates of additive/multiplicative transport costs (TC) \hyperlink{app_compeffects_strategy}{\beamergotobutton{Comparison with Hummels (2007)}}\vspace{0.1cm}
\begin{itemize}
\item[-] That vary over time ($t$)/product ($k$) /origin country ($i$) (by transport mode): $\widehat{\tau}_{ikt}, \widehat{t}_{ikt}$ \vspace{0.1cm}
\end{itemize}
\item Starting from these values, extract the ``pure'' TC component \hyperlink{app_compeffects_details1}{\beamergotobutton{More details}}\vspace{0.1cm}
\begin{itemize}
\item[-] By the mean of a time fixed-effect \vspace{0.1cm}
\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted \vspace{0.2cm}
%\item[-] From which we extract the fitted ``total'' transport cost measure \vspace{0.1cm}
\end{itemize}
\item What explains the decrease of transport costs over time?  \vspace{0.1cm}
\begin{itemize}
\item[-] Composition effects vs ``Pure'' TC changes? \vspace{0.1cm}
%\item[-] Is it mostly due to the composition of trade (product quality/weight, less distant countries of origin...) \vspace{0.1cm}
%\item[-] or to the ``pure'' transports costs dimension (e.g., handling costs) ?
\end{itemize}
\item[$\Rightarrow$] Compare the unfitted TC component to the fitted one (composition effects excluded) \vspace{0.1cm}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Time trends in transport costs: Comments}
\begin{itemize}
\item A strong reduction of transport costs over the period (1974-2013) \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
\item[-] Consistent with the literature, see Lafourcade \& Thisse (2011) \vspace{0.1cm}
\item[-] Roughly same order of magnitude for both the ad-valorem and the additive component  \vspace{0.1cm}
\normalsize
\end{itemize}
\item \textbf{Trade composition effects: A minor role in accounting for the time trend of TC}\vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] Not much evidence of a difference between the ``raw'' and the ``pure'' TC components \vspace{0.1cm}
\normalsize
\end{itemize}
\item In Air transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 50\% decrease, mainly attributable to the reduction of ``pure'' transport costs \vspace{0.1cm}
\item[-] For both the additive and the ad-valorem components  \vspace{0.1cm}
\end{itemize}
\normalsize
\item Composition effects (slightly) more pronounced in maritime transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 60\% decrease: Roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
\item[-] But for the multiplicative component only \vspace{0.1cm}
\end{itemize}
\normalsize
\end{itemize}
\end{frame}

\begin{frame}[label = slide_compeffects_results]
\frametitle{Time trends in transport costs: The role of additive costs}
\begin{itemize}
\item A result in sharp contrast with Hummels (2007) \vspace{0.1cm}
\begin{itemize}
\item[-] According to whom pure transport costs have reduced much more than the unfitted ones (over 1974-2004) \vspace{0.1cm}
\item[-] Suggesting an important role to trade composition effects  \vspace{0.1cm}
\end{itemize}
\item Not the case here. \vspace{0.1cm}
\item[$\Rightarrow$] \textbf{Where does the difference come from?} \vspace{0.1cm}
\item \textbf{From the treatment of the additive component of transport costs} \vspace{0.1cm}
\begin{itemize}
\item[-] Assuming (as we do, $\neq$ Hummels (2007)) a varying share of the additive component over time, product and country partner is key \vspace{0.1cm}
\item[-] Strong impact in the decomposition of the time trend of transport costs, between composition effects and ``pure'' transport costs changes \vspace{0.1cm}
\item[-] See  \hyperlink{app_compeffects_results}{\beamergotobutton{Figure 3}}
\end{itemize}\vspace{0.1cm}
\item[$\Rightarrow$] The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs
\end{itemize}

@@

>>>>

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] Part of the ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] With $p$ the import price, $\widetilde{p}$ the export price, $q$ the quantity traded and $\tau >1$ the trade cost \vspace{0.1cm}
\item[$\Rightarrow$] Multiplicative costs:     \footnotesize $pq = \tau \widetilde{p}q$
\normalsize
\end{itemize}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Would not trade costs rather exhibit an additive structure ?  \vspace{0.1cm} \vspace{0.1cm}
\begin{itemize}
\item[-] Why would it be more costly to transport from Milan to Paris, a pair of Italian shoes at price \euro 300 than a pair of Italian shoes at price \euro 50? \vspace{0.1cm}
\item[-] Pricing shipping often includes an additive component \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] UPS: A \$125 fee charged for a 2 pound package from Oslo to NYC (Irrarazabal et al., 2015) \vspace{0.1cm}
%\item[$\ast$] Some trade policy instruments as well (quota license price) \vspace{0.1cm}
\item[$\Rightarrow$] Additive costs: \footnotesize $pq = (t+\widetilde{p})q  $
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item Recent empirical evidence in support of the additive structure of trade costs \vspace{0.1cm}
\begin{itemize}
\item[-] Irrarazabal et al. (2015), Hummels \& Skiba (2004), Martin (2012)  \vspace{0.2cm}
\end{itemize}
\item The structure (additive vs iceberg) of transport costs is not anecdotal \vspace{0.1cm}
\begin{itemize}
\item[-] Additive costs play an important role in shaping the pattern of trade flows (Alchian \& Allen, 1964) \vspace{0.1cm}
\item[-] Strong normative implications, notably w.r.t. the welfare gains of trade liberalization (Sorensen, 2014)\vspace{0.1cm}

\end{itemize}
\item[$\Rightarrow$] Transport costs are likely to display an additive component, but precisely... by how much? \vspace{0.2cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}

@@

%}
\end{itemize}
\item[(2)] \textbf{In terms of quality of fit}: With the additive component included, \vspace{0.1cm}
\begin{itemize}
%\footnotesize{
\item[-] The estimated iceberg component is reduced by a factor of 2 \vspace{0.1cm}
\item[-] A substantially better ``goodness-of-fit'' %}
%}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item[(3)] \textbf{In accounting for the time trend of transport costs} \vspace{0.2cm}
\begin{itemize}
\item[-] The importance of modelling the additive component of transport costs \vspace{0.1cm}
\item[-] Because it changes the result of excluding the trade composition effects when characterizing the time trend \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] The decrease in TC over time is mostly attributable to the decrease in the ``pure'' transport costs,  \vspace{0.1cm}
\item[$\ast$] Rather than to trade composition effects  \vspace{0.1cm}
\end{itemize}
\item[-] A result in sharp contrast with Hummels (2007) \vspace{0.2cm}
\end{itemize}
\end{itemize}
\textbf{All our results: Provide new quantitative evidence about the importance of the additive component in international transport costs}

@@

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the ``export'' price and the ``import'' price of US imports\vspace{0.1cm}
\item Database: US Imports of Merchandise database \vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, SITC 5-digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
%(Could go up to HS-10 with recent data)\vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: Based on the ratio $p/\widetilde{p}$ \vspace{0.1cm}
\item The RHS is at the 3-digit classification level  \vspace{0.1cm}
\begin{itemize}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 sectors (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 sectors at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{More on our database}
\begin{itemize}
\item Implications (and limitations) \vspace{0.1cm}
\begin{itemize}
\item[-] Only cover international transport costs \vspace{0.1cm}
\item[-] Among transport costs, insurance + handling + quantitative freight costs (e.g. not those related to the time value of goods) \vspace{0.1cm}
%\item[$\Leftrightarrow$] Among the 21\% markup from AVW, $\simeq$ 11\% for freight costs (the remaining 9\% for time costs). But ``rough'' estimate. Precisely?  \vspace{0.1cm}
\end{itemize}
\item A rich database to exploit \vspace{0.1cm}
\begin{itemize}
\item[-] US imports, long time period: Broad view of international trade flows \vspace{0.1cm}
\item[-] A reliable database, already used by Hummels (2007), but on a shorter length of time \vspace{0.1cm}
\item[-] Have both the import and the export prices  \vspace{0.1cm}
\item[$\Rightarrow$] Allows the estimate the levels of both the ad-valorem and the additive transport costs  \vspace{0.1cm}
\begin{itemize}
\item[$\neq$] Irarrazabal et al. (2015)
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The equation at the root of the estimation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For a product $k$, from country $i$ (for a given transport mode)   \vspace{0.1cm}
\item Rewrite to get:
\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:basis_equation}
\end{equation}
\item[$\Rightarrow$] For each year over 1974-2013 (at the $k=$ 5-digit classification level) \vspace{0.1cm}
\begin{itemize}
\item[-] The equation is also mode (air or vessel) - specific
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label = slide_empirical_strategy]
\frametitle{Empirical specification (2)}
\begin{itemize}
\item Make some simplifying assumptions \hyperlink{app_empirical_strategy}{\beamergotobutton{More details}}  \vspace{0.1cm}
\begin{itemize}
\item[-] About the specification of both additive and multiplicative costs  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Separability between the product and the country dimensions (as Irarrazabal et al., 2015)   \vspace{0.1cm}
\item[$\ast$] All products $k$ in a 3-digit sector ($s$) share the same structure of costs  \vspace{0.1cm}
\end{itemize}
\item[-] About the specification of the error term \vspace{0.1cm}
\end{itemize}
\item Taking logs, we finally estimate the following equation:
\footnotesize
\begin{equation*}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik}
\end{equation*}

\begin{itemize}
\footnotesize
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0  \vspace{0.1cm}
\item[-] $\tau_i$, $\tau_{s(k)}$, $t_i$, $t_{s(k)}$ are the parameters (i.e. fixed effects) to be estimated
\end{itemize}
\normalsize
\item A non-linear equation (due to the additive costs)  
\item[$\Rightarrow$] Estimation using non-linear squares 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (3)}
\textbf{How to characterize the importance of additive costs relatively to iceberg?} \vspace{0.1cm}
\begin{itemize}
\item Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\item[$\Rightarrow$] Estimate \textit{two} models\vspace{0.1cm}
%\begin{itemize}
\item[(A)] With only ad-valorem costs:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\normalsize
\item[(B)] With additive costs included:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\begin{itemize}
\item[$\ast$] Under Model (B), $\simeq$ 800 fixed effects to estimate  \vspace{0.1cm}
\item[$\ast$] For each year over 1974-2013, by transport mode (air/vessel)
%\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_result1]
\frametitle{Result 1: Size of additive transport costs \hyperlink{app_result1}{\beamergotobutton{More on this}} }
\begin{itemize}
\item \textbf{International transport costs: A sizeable additive component}  \vspace{0.2cm}
\begin{itemize}
\item[-] 48.2\% of total costs in average for vessel, 42.3\% for air  \vspace{0.1cm}
\item[-] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1}} \vspace{0.1cm}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.2cm}
\begin{itemize}
\footnotesize{
\item[$\ast$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\ast$] From 5.8\% to 3.2\% in maritime transport (mean value over the period) \vspace{0.1cm}
\item[$\ast$] From 5.1\% to 2.5\% in Air transport }\vspace{0.1cm}
\end{itemize}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp.  
\item[-] Additive cost: 1.8\% and 2.9\% of the export price \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Valuable insights for related papers more theoretically-oriented, in need for calibration
\end{itemize}
\end{frame}

@@

\begin{frame} [label = slide_compeffects_strategy]
\frametitle{Time trends in transport costs \& composition effects }
\begin{itemize}
\item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.2cm}
\item The change in overall transport costs over time: \vspace{0.1cm}
\begin{itemize}
\item[-] Depend on the evolution of per product- per partner costs, \vspace{0.1cm}
\item[-] But also on the composition of trade flows \vspace{0.1cm}
\begin{itemize}
\item[$\star$] Over time, import more goods that are cheaper to transport, and/or from countries with which it is cheaper to trade \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) \vspace{0.1cm}
\item Some insights on our methodology \hyperlink{app_compeffects_strategy}{\beamergotobutton{More details}} \hyperlink{app_compeffects_Hummels}{\beamergotobutton{Comparison with Hummels (2007)}} \vspace{0.1cm}
\begin{itemize}
\item[-] Start with estimates of additive/multiplicative transport costs  \vspace{0.1cm}
\item[-] Extract the ``pure'' TC component by the mean of a time fixed-effect \vspace{0.1cm}
%\item[-] For both the additive \& the multiplicative TC components \vspace{0.1cm}
\item[-] Build a unified measure of ``total'' transport costs (agglomerate the two components), also both fitted and unfitted
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Time trends in transport costs: Comments}
\begin{itemize}
\item A strong reduction of transport costs over the period (1974-2013) \vspace{0.1cm}
\begin{itemize}
\footnotesize
%\item[-] By 50\% in Air and 60\% in Ocean transport \vspace{0.1cm}
\item[-] Consistent with the literature, see Lafourcade \& Thisse (2011) \vspace{0.1cm}
\item[-] Roughly same order of magnitude for both the ad-valorem and the additive component  \vspace{0.1cm}
\normalsize
\end{itemize}
\item \textbf{Trade composition effects: A minor role in accounting for the time trend of TC}\vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] Not much evidence of a difference between the ``raw'' and the ``pure'' TC components \vspace{0.1cm}
\normalsize
\end{itemize}
\item In Air transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 50\% decrease is mainly attributable to the reduction of ``pure'' transport costs, for both components  \vspace{0.1cm}
\end{itemize}
\normalsize
\item Composition effects (slightly) more pronounced in maritime transport \vspace{0.1cm}
\begin{itemize}
\footnotesize
\item[-] The 60\% decrease is roughly 50\% decrease of ``pure'' TC, 10\% decrease from composition effects \vspace{0.1cm}
\item[-] But for the multiplicative component only \vspace{0.1cm}
\end{itemize}
\normalsize
\end{itemize}
\end{frame}

\begin{frame}[label = slide_compeffects_results]
\frametitle{Time trends in transport costs: The role of additive costs}
\begin{itemize}
\item A result in sharp contrast with Hummels (2007) \vspace{0.1cm}
\begin{itemize}
\item[-] According to whom pure transport costs have decreased much more than the unfitted ones (over 1974-2004) \vspace{0.1cm}
\item[-] Suggesting an important role to trade composition effects  \vspace{0.1cm}
\end{itemize}
\item Not the case here. \vspace{0.1cm}
\item[$\Rightarrow$] \textbf{Where does the difference come from?} \vspace{0.1cm}
\item \textbf{From the treatment of the additive component of transport costs} \vspace{0.1cm}
\begin{itemize}
\item[-] Assuming (as we do, $\neq$ Hummels (2007)) a varying share of the additive component over time, product and country partner is key \vspace{0.1cm}
%\item[-] Strong impact in the decomposition of the time trend of transport costs, between composition effects and ``pure'' transport costs changes \vspace{0.1cm}
\item[-] What if we relax this assumption?
\end{itemize}\vspace{0.1cm}
\end{itemize}

@@

\begin{frame}
%\frametitle{Time trend of transport costs \& the modeling of the additive component}
\begin{itemize}
\item Assuming (as Hummels, 2007) a constant share of the additive component over time/country partner/product \vspace{0.1cm}
\item[$\Rightarrow$] Transport cost changes decompose as:
\begin{figure}[htbp]
%\caption{Transport costs (with and without composition effects)}
%\label{fig:totalTC_compeffects_excl}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Air } & {\scriptsize (b) Vessel}\\
\includegraphics[width=5cm, height=4cm]{figure5_comme_hummels_air.pdf}
& \includegraphics[width=5cm,height=4cm]{figure6_comme_hummels_ocean.pdf} \\
\end{tabular}
\end{center}
\end{figure}
\item[$\Rightarrow$] Suggest a much more important role to trade composition effects...  \vspace{0.1cm}
\item Than they actually do, see \hyperlink{slide_compeffects_figure}{\beamergotobutton{Figure 2}} \vspace{0.1cm}

\item[$\Rightarrow$] The (appropriate) modeling of additive costs, of key importance in understanding the time trends of international transport costs
\end{itemize}
\end{frame}
